---
title: "Evolutioary trees"
author: "Julieanna"
format: html
editor: visual
---

## Description

In this session we will be learning how to construct phylogeny trees in R studio. We will be looking at different ways to visualize the data in an appropriate way depending on the research study.

## Mammal tree

The phylogeny tree seems is about mammals and how they are related to other species. It makes sense that apes are the most common species to us but I thought it was interesting that they included Bovine and rodents as out groups as even though they are mammals as they look nothing like the other groups.

## Visualizing trees

```{r}
ggtree(itol)
```

This is a basic out line of the tree of life.

```{r}
ggtree(itol) +
  geom_tiplab(color = "blue", size = 1)
```

This is the same basic tree of life but with labels, It is hard to read like this.

```{r}
ggtree(itol, layout = "circular") +
  geom_tiplab(color = "blue", size = 2)
```

This circular graph is easier to read and visualize.

```{r}
ggtree(itol, layout = "circular") +
  geom_tiplab(color = "blue", size = 2) +
  geom_strip(13, 14, color = "red", barsize = 1)


```

This version has a specific section of the tree highlighted.

```{r}
ggtree(itol) +
  coord_flip() +
  scale_x_reverse()

```

This version is better for visualization and seeing the difference in time between species.

```{r}
ggtree(itol, layout = "unrooted")
```

This is an unrooted tree that is used when comparing species without the component of time.

```{r}
ggtree(mammal) +
  geom_tiplab(color = "blue", size = 2)
```

A basic phylogeny tree of mammals looking from the lenses of humans

```{r}
ggtree(mammal, layout = "circular") +
  geom_tiplab(color = "blue", size = 2)
```

This is a basic circular tree of mammals.

## Conclusion

In this section we learned how to create phylogeny trees and different ways of visualizing them. I am curious how an unrooted tree would be used in practice as I feel like I would need more experience to understand it truly even if I understand conceptually.
