---
title: "RNA seq 2"
author: "Julieanna"
format: html
editor: visual
---

## Description

Using R studio we are creating a heat map of the plant species *Arabidopsis thaliana* and see how gene expression differs between different parts of the plant and different genetic variations of the plant.

```{r}
(!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("ComplexHeatmap")
```

```{r}
mat <- log(as.matrix(at_tf_gex[ , 5:55]))
ecotype <- stringr::str_split(colnames(mat), ",", simplify = TRUE)[,1]
part <- stringr::str_split(colnames(mat), ",", simplify = TRUE)[,2]

```

```{r}
top_annot <- HeatmapAnnotation("Ecotype" = ecotype, "Plant Part" = part, col = list("Ecotype" = ecotype_colors, "Plant Part" = part_colors), annotation_name_side = "left")

side_annot <- rowAnnotation(length = anno_points(at_tf_gex$Length, pch = 16, size = unit(1, "mm"), axis_param = list(at = seq(1, max(at_tf_gex$Length), length.out = 4)),))
```

```{r}
ht_1 <- Heatmap(mat, name="log(TPM)", row_km = 6, col = data_col_func, top_annotation = top_annot, right_annotation = side_annot, cluster_columns = TRUE, column_split = ecotype, show_column_names = FALSE, column_title = " ")

ComplexHeatmap::draw(ht_1)
```

The ecotype represents different genetic variations of the *Arabidopsis thaliana* and the plant type represents different part of the plants. The log(TPM) shows gene expression meaning that the lighter the color is the more highly expressed that gene is. In the graph you can see that genes are not as expressed in column 3 in the root part of the plant. similarly in column 2 only the roots are highly expressed.
